\textbf{Advection versus diffusion in intraventricular transport}

### 2025-03-24 paper/main.tex
methods and discussion updates

\author[1,*]{Halvor Herlyng}
\author[1]{Ada J.~Ellingsrud}
\author[1]{Miroslav Kuchta} 
\author[2]{Inyoung Jeong}
\author[1,3]{Marie E.~Rognes}
\author[2, *]{Nathalie Jurisch-Yaksi}
\affil[1]{Department for Numerical Analysis and Scientific Computing, Simula Research Laboratory, Oslo, Norway} % Kristian Augusts gate 23, 0164
\affil[2]{Department of Clinical and Molecular Medicine, Norwegian University of Science and Technology, Trondheim, Norway} % Erling Skjalgsons gate 1, 7491
\affil[3]{K.~G.~Jebsen Center for Brain Fluid Research, Oslo, Norway}
\affil[*]{\href{mailto:hherlyng@simula.no}{hherlyng@simula.no}, \href{mailto:nathalie.jurisch-yaksi@ntnu.no}{nathalie.jurisch-yaksi@ntnu.no}}
\date{}

@@

\begin{figure}%[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure2_simulation_results_different_coeffs.png}\caption{Simulated distribution and evolution after photoconversion of
    extracellular vesicles (EVs) ($D_1$), STM-GFP ($D_2$) and Dendra2 ($D_3$). 
    \textbf{a.} Geometry-centered clip of the ventricles mesh showing the locations of the ROIs.
    \textbf{b--d.} Slices of the concentration field simulated
    with diffusion coefficients $D_1$, $D_2$ and $D_3$ for nine time instants covering the
    time range from 90 seconds to 811 seconds, corresponding to 200 to 1800 cardiac cycles,
    for even intervals of 200 cycles.
    Slices are taken in the $xz$-plane ($y=0.140$ mm, the center of the geometry).
    The colorbar is scaled so the maximum is the maximum concentration of the Dendra2 
    concentration field slice at the last time instant.
    \textbf{e--g.} Mean concentration $\cbar$ in each region of interest (ROI)
    as function of time for diffusion coefficients $D_1$, $D_2$ and $D_3$, respectively. 
    \textbf{h.} The time $\hat{t}$ before the mean concentration $\cbar$ exceeds a
    threshold value of $0.25$ (ROIs 2-4) or $0.10$ (ROIs 5, 6) in each simulation setup.
    Note that there is no bar for $D_1$ in ROI 6, because the threshold value was never exceeded.}
    \label{fig:fig2}
\end{figure}

@@

Comparing the intensity changes with the simulated concentration
profiles in ROIs 1--6 (\Crefpanels{fig:fig3}{a, b}) show that the
model predictions agree well with the experimental measurements. We
note minor differences however. The simulated values of $\cbar_i$ at
the final time \fixme{(0.72, 0.82, 0.56, 0.35 and X in ROIs 2--6, respectively)}
are generally higher than the corresponding experimental $\Delta F$
values (0.63, 0.71, 0.67, 0.29 and 0.18). Moreover, we observe
differences in the distribution within the middle ventricle,
especially with regard to ROIs 2 and 3. Overall, we note that the time
to reach a 25\% threshold agree well between the model predictions and
experimental values (\Crefpanels{fig:fig3}{c, d}).

@@

\subsection{Modification of ventricular geometry impacts solute distribution}
Variations in ventricular morphology impact CSF flow patterns~\cite{Olstad2019CiliaryDevelopment}.
To assess how these variations influence solute distribution,
we consider four alternative brain ventricles meshes:
one where the duct connecting the anterior and middle ventricles is constricted
(66 \% reduction in cross-sectional area);
one where the duct connecting the middle and posterior ventricles is constricted
(33 \% reduction);
one where the middle ventricle is shrunk in the lateral direction (43 \% reduction);
and the combination of the above three modifications (\Crefpanels{fig:fig6}{a, b}). In these
four modified geometries the volume is reduced by 3.9\%, 4.1\%, 1.4\% and 9.4\%, respectively,
compared to the original geometry.

@@

With an interest in assessing the impact of ventricular morphology on solute transport,
we simulated transport on modified ventricular geometries.
Notably, shrinking ducts reduced interventricular solute movement, while shrinking
the middle ventricle reduced diffusive transport upstream of the flow field.
We previously identified large inter-individual variation in ventricular size and
CSF velocities between larva at given developmental stages and throughout
development~\cite{Olstad2019CiliaryDevelopment, DGama2025MotileBrain, Jeong2024TheZebrafish}.
Our results suggest that these two parameters, CSF velocities and ventricular size,
may well co-vary and explain the variation in the two.

@@

\subsection{Computational modeling aspects}
We chose to model the cilia as a force
exerted tangentially on the boundary, as opposed to a
couple of previous continuum model simulation studies of
cilia-mediated flow where cilia were represented by body
forces,~\cite{Siyahhan2014FlowVentricles, Thouvenin2020OriginCanal}.
We employed slip boundary conditions at the ventricular walls,
motivated by experimental observations of the
flow field in zebrafish embryo brain ventricles, both previously and in this work, in which
the fluid velocity increases towards the
wall~(\Crefpanels{fig:fig1}c)~\cite{Olstad2019CiliaryDevelopment}. 
In addition to its favorable mass-conserving property, the 
Brezzi-Douglas-Marini and discontinuous Galerkin (BDM-DG) discretization scheme 
we employed for the Stokes equations
lets us impose the impermeability condition strongly, and the tangential
traction boundary condition weakly, without any tailored approach,
since these boundary conditions appear naturally in the weak form of the
Stokes equations~(\Cref{subsec:appendixA1}). If instead e.g. Taylor-Hood
elements were to be used, the impermeability boundary condition must
be handled with for example a Nitsche method~\cite{Nitsche1971UberSind}
or with a Lagrange multiplier~\cite{Babuska1973TheMultipliers,
Bertoluzza2017BoundaryHemodynamics}.

The BDM-DG scheme ($\mathrm{BDM}_1$-$\mathrm{DG}_0$) that we employed for the Stokes equations
is computationally more expensive than a more conventional approach based on
e.g. Taylor-Hood discretization~\cite{Stenberg1990ErrorProblem}
($P_2-P_1$ elements) or lower-order enriched
elements~\cite{Brezzi2011MixedMethods},
such as the mini-element ($P_1-P_1$ with a velocity stabilization).
Here, $P_k$ is the space of continuous Lagrange polynomials of degree less
than or equal to $k$.
We opted for the BDM-DG scheme because of its exactly-mass-conserving property,
as the elements by definition satisfy the incompressibility constraint for the fluid.
This leads to $\normlinf{\nabla\cdot\uu_h}$ being equal to zero to within solver tolerances,
an important property for the stability of the advection-diffusion equation\cite{CITATION}.

@@

\subsection{Image-based computational geometries of zebrafish brain ventricles}
To model CSF flow and transport in the zebrafish brain ventricles, we
generated a 3D surface representation of the ventricular walls using
confocal imaging data of a zebrafish embryo injected
intraventricularly with a 70 kDa dye at 2
dpf~\cite{Olstad2019CiliaryDevelopment}. At 2 dpf, the ventricular
system consists mainly of 3 cavities: the telencephalic (anterior),
the di-/mesencephalic (middle) and the rhombencephalic (posterior)
ventricles, which are connected by interventricular ducts. From the
surface geometry, we constructed a volumetric mesh of the ventricles
by initially using fTetWild~\cite{Hu2020FastWild} and subsequently
local refinement functionality in DOLFINx~\cite{dolfinx2023preprint}. The standard mesh
consists of 141 512 tetrahedral cells with a maximal (minimal) edge length of 14.9 \textmu
m (3.1 \textmu m) (\Crefpanels{fig:fig1}{d}). Parts of the
ventricular surfaces were marked to be lined with motile cilia
(\Crefpanels{fig:fig1}{d}, magenta
markers)~\cite{Olstad2019CiliaryDevelopment}.

@@

\subsection{Computational CSF dynamics in the brain ventricles}
The beating motion of the cilia generates CSF flow with persistent rotational
structures~\cite{Olstad2019CiliaryDevelopment}
within the brain ventricles (\Crefpanels{fig:fig1}{a--c}).
In addition, cardiac pulsations induce pulsatile CSF flow.
We model this flow of CSF by the time-dependent,
incompressible Stokes equations which read as follows.
Find the CSF velocity $\uu = \uu(\xx, t) = (u_x(\xx, t), u_y(\xx, t), u_z(\xx, t))$
and the CSF pressure $p = p(\xx, t)$ for $\xx \in \Omega$ and time $t>0$, such that
\begin{subequations}
    \begin{align}
      \rho \pdifft{\uu} - \nabla \cdot \bsig &= \mathbf{0}
      \quad \text{in } \Omega,\\
      \nabla \cdot \uu &= 0
      \quad  \text{in } \Omega,
  \end{align}
  \label{eq:stokes_eqs}% 
\end{subequations}%
where the stress tensor is defined as $\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$
and $\beps(\uu) = \frac{1}{2}\left(\nabla \uu + (\nabla\uu)^T\right)$ is the strain-rate tensor. 
Bold-face characters denote vectors or tensors, and $\mathbf{I}$ is the identity tensor in three dimensions.
Furthermore, $\rho=1000 \ \mathrm{kg/m^3}$ is the CSF density
and $\mu=0.7 \ \mathrm{mPa\cdot s}$ is the dynamic viscosity
of the CSF~\cite{Bloomfield1998EffectsFluid}.
We introduce the tangential traction $\bsigpar$ as the
tangential component of the traction $\hat{\bsig}=\bsig\nn$:
\begin{equation*}
    \bsigpar = P_{\nn}(\hat{\bsig}) = (\mathbf{I} - \nn\otimes\nn)\hat{\bsig},
\end{equation*}
where $\nn$ is the outer unit normal of the boundary surface $\Gamma = \partial\Omega$,
and $P_{\nn}(\rr)$ is the tangential projection of a vector $\rr$ onto $\Gamma$.
We consider three types of boundary conditions:
(i) \emph{tangential traction} (no flow normal to the boundary
$\uu \cdot \nn = 0$ with a stress $\btau$ applied tangentially on the boundary $\bsigpar = \btau$),
(ii) \emph{normal pressure} (a pressure $\Tilde{p}$ applied in
the normal direction: $(\mu\nabla\uu - p\mathbf{I})\nn = \Tilde{p}(t)\nn$),
and (iii) \emph{free slip} (no flow normal to the boundary
$\uu \cdot \nn = 0$ and no tangential traction $\bsigpar = {\bf 0}$).
Unless noted otherwise, the free-slip condition is applied on the
boundary. We assume that the system starts at rest: $\uu(\xx, t=0) = \mathbf{0}$.

To study the impact of the motile cilia and the cardiac pulsations on the CSF flow,
we consider the following computational flow scenarios,
each predicting a CSF flow velocity $\uu$ and pressure $p$.
\begin{itemize}
    \item A \emph{baseline} flow model representing a wild-type zebrafish including 
    contributions both from the cilia forces and the cardiac pulsations.
    In this scenario, we apply the tangential traction $\btau$ on the cilia boundary
    $\Gc$ and prescribe the normal pressure $\Tilde{p}$ on the anterior/posterior boundary $\Gp$. % Model 0/Model C

    \item A \emph{cilia-driven/no-cardiac} flow model including contributions from the
    cilia forces but no cardiac pulsations.
    Here, we apply a tangential traction $\btau$ on the cilia boundary $\Gc$.
    In the absence of cardiac flow, the anterior/posterior boundary $\Gp=\emptyset$. % Model I/Model A

    \item A \emph{cardiac-induced/no-cilia} flow model including contributions from the
    cardiac pulsations but no cilia forces.
    Here, we set the tangential traction to zero on the cilia boundary
    $\Gc$, while prescribing the normal pressure $\Tilde{p}$ on the
    anterior/posterior boundary $\Gp$.  % Model II/Model B
\end{itemize}
The tangential traction and normal pressure are described further below.

@@

\subsection{Computational model of solute transport after photoconversion}
We simulate transport of photoconverted proteins within the brain ventricles by
modeling transport of a solute with concentration $c(\xx, t)$ for $\xx\in\Omega$ and
time $t > 0$ by the advection-diffusion equation
\begin{equation}
    \pdifft{c} + \nabla\cdot\JJ = 0 \quad \text{in } \Omega,
    \label{eq:adv_diff_strong}
\end{equation}
where the total concentration flux $\JJ$ is assumed to consist of an advective and a diffusive flux:
\begin{equation*}
    \JJ = c\uu - D\nabla c .
\end{equation*}
Here, $\uu$ is the CSF velocity field, governed by the Stokes equations (\Cref{eq:stokes_eqs}),
and $D$ is the molecular diffusion coefficient of the solute. 
The ventricular walls are assumed to be impermeable (admitting \emph{no flux}: $\JJ\cdot\nn=0$),
except at the anterior/posterior boundary $\Gp$.
At the anterior/posterior boundary, we permit the solute concentration to be advected into
and out of the domain in a manner that ensures conservation of mass
(see~\Cref{subsec:adv_diff_eq_weak_form} for details).

@@

The numerical solution schemes for the Stokes equations and the
advection-diffusion equation were implemented and solved numerically with
DOLFINx~\cite{dolfinx2023preprint} with petsc4py~\cite{Dalcin2011ParallelPython}
as linear algebra backend. The simulation code is openly
available\fixme{~\cite{zenodo-link-to-all-code}}.
A direct solver using MUMPS~\cite{Amestoy2011Mumps} was used to solve the
linear systems resulting from the discretized Stokes equations.
For the discretized advection-diffusion equation, we used an iterative
linear solver based on GMRES~\cite{Saad1986GMRES:Systems} with block Jacobi
preconditioning~\cite{Jacobi1845UeberGleichungen}. We discretize the time-dependent 
diffusion equation (without advection) and use the assembled linear system matrix
as the preconditioner matrix for the advection-diffusion problem.
Numerical experiments verified the accuracy of the model implementation
(\Cref{sec:appB_verification}).

>>>>

\author[1,*]{Halvor Herlyng}
\author[1]{Ada J.~Ellingsrud}
\author[1]{Miroslav Kuchta} 
\author[2]{Inyoung Jeong}
\author[1,3]{Marie E.~Rognes}
\author[2, *]{Nathalie Jurisch-Yaksi}
\affil[1]{Department for Numerical Analysis and Scientific Computing, Simula Research Laboratory, Oslo, Norway} % Kristian Augusts gate 23, 0164
\affil[2]{Department of Clinical and Molecular Medicine, Norwegian University of Science and Technology, Trondheim, Norway} % Erling Skjalgsons gate 1, 7491
\affil[3]{K.~G.~Jebsen Center for Brain Fluid Research, Oslo, Norway}
\affil[*]{\href{mailto:hherlyng@simula.no}{hherlyng@simula.no}, \href{mailto:nathalie.jurisch-yaksi@ntnu.no}{nathalie.jurisch-yaksi@ntnu.no}}
\date{\vspace{-2em}}

@@

\begin{figure}%[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure2_simulation_results_different_coeffs.png}
    \caption{Simulated distribution and evolution after photoconversion of
    extracellular vesicles (EVs, $D_1$), STM-GFP ($D_2$) and Dendra2 ($D_3$). 
    \textbf{a.} Geometry-centered clip of the ventricles mesh showing the locations of the 
    regions of interest (ROIs).
    \textbf{b--d.} Slices of the concentration field simulated
    with diffusion coefficients $D_1$, $D_2$ and $D_3$ for nine time instants covering the
    time range from 90 seconds to 811 seconds, corresponding to 200 to 1800 cardiac cycles,
    for even intervals of 200 cycles.
    Slices are taken in the $xz$-plane ($y=0.140$ mm, the center of the geometry).
    \textbf{e--g.} Mean concentration $\cbar$ in each ROI
    as function of time for diffusion coefficients $D_1$, $D_2$ and $D_3$. 
    \textbf{h.} The time $\hat{t}$ before the mean concentration $\cbar$ exceeds a
    threshold value of $0.25$ (ROIs 2-4) or $0.10$ (ROIs 5, 6) in each simulation setup.
    Note that there is no bar for $D_1$ in ROI 6, because the threshold value was never exceeded.}
    \label{fig:fig2}
\end{figure}

@@

Comparing the intensity changes with the simulated concentration
profiles in ROIs 1--6 (\Crefpanels{fig:fig3}{a, b}) show that the
model predictions agree well with the experimental measurements. We
note minor differences however. The simulated values of $\cbar_i$ at
the final time (0.80, 0.76, 0.86, 0.60 and 0.38 in ROIs 2--6, respectively)
are generally higher than the corresponding experimental $\Delta F$
values (0.63, 0.71, 0.67, 0.29 and 0.18). Moreover, we observe
differences in the distribution within the middle ventricle,
especially with regard to ROIs 2 and 3. Overall, we note that the time
to reach a 25\% threshold agree well between the model predictions and
experimental values (\Crefpanels{fig:fig3}{c, d}).

@@

\subsection{Modification of ventricular geometry impacts solute distribution}
Variations in ventricular morphology impact CSF flow patterns~\cite{Olstad2019CiliaryDevelopment}.
To assess how these variations influence solute distribution,
we consider four alternative ventricular geometries:
one where the duct connecting the anterior and middle ventricles is constricted
(66 \% reduction in cross-sectional area);
one where the duct connecting the middle and posterior ventricles is constricted
(33 \% reduction);
one where the middle ventricle is shrunk in the lateral direction (43 \% reduction);
and the combination of the above three modifications (\Crefpanels{fig:fig6}{a, b}). In these
four modified geometries the volume is reduced by 3.9\%, 4.1\%, 1.4\% and 9.4\%, respectively,
compared to the original geometry.

@@

With an interest in assessing the impact of ventricular morphology on solute transport,
we simulated transport in modified ventricular geometries.
Notably, shrinking ducts reduced interventricular solute distribution, while shrinking
the middle ventricle reduced diffusive transport upstream of the flow field.
We previously identified large interindividual variation in ventricular size and
CSF velocities between larva at given developmental stages and throughout
development~\cite{Olstad2019CiliaryDevelopment, DGama2025MotileBrain, Jeong2024TheZebrafish}.
Our results suggest that these two parameters, CSF velocities and ventricular size,
may well co-vary and explain the variation in the two.

@@

\subsection{Computational modeling aspects}
We modeled motile cilia as a stress
exerted tangentially on the CSF at the ventricular walls
while employing a slip boundary condition,
motivated by experimental observations of the
flow field in zebrafish embryo brain ventricles, both previously and in this work, in which
the fluid velocity increases towards the
wall~(\Crefpanels{fig:fig1}c)~\cite{Olstad2019CiliaryDevelopment}. 
The BDM-DG scheme ($\mathrm{BDM}_1$-$\mathrm{DG}_0$) we used to discretize the
governing equations of the flow is computationally more expensive
than more conventional approaches such as a
Taylor-Hood discretization
($P_2-P_1$ elements)~\cite{Stenberg1990ErrorProblem}, or lower-order enriched
elements, such as the mini-element
($P_1-P_1$ with a velocity stabilization)~\cite{Brezzi2011MixedMethods}.
Here, $P_k$ is the space of continuous Lagrange polynomials of degree less
than or equal to $k$.
We opted for the BDM-DG scheme because of its exactly-mass-conserving property;
by the definition of the elements, the numerical approximation of
the CSF velocity satisfies the continuity equation (\Cref{eq:stokes_eqs_cont}).
This leads to $\normlinf{\nabla\cdot\uu_h}$ being equal to zero to within solver tolerances,
an important property for the stability of the advection-diffusion equation\cite{CITATION}.
In addition to its favorable mass-conserving property, the 
Brezzi-Douglas-Marini and discontinuous Galerkin (BDM-DG) discretization scheme 
we employed for the Stokes equations
lets us impose the impermeability condition strongly, and the tangential
traction boundary condition weakly, without any tailored approach,
since these boundary conditions appear naturally in the weak form of the
Stokes equations~(\Cref{subsec:appendixA1}). If instead e.g. Taylor-Hood
elements were to be used, the impermeability boundary condition must
be handled with for example a Nitsche method~\cite{Nitsche1971UberSind}
or with a Lagrange multiplier~\cite{Babuska1973TheMultipliers,
Bertoluzza2017BoundaryHemodynamics}.

@@

\subsection{Image-based computational geometries of zebrafish brain ventricles}
To model CSF flow and transport in the zebrafish brain ventricles, we
generated a 3D surface representation of the ventricular walls using
confocal imaging data of a zebrafish embryo injected
intraventricularly with a 70 kDa dye at 2
dpf~\cite{Olstad2019CiliaryDevelopment}. At 2 dpf, the ventricular
system consists mainly of 3 cavities: the telencephalic (anterior),
the di-/mesencephalic (middle) and the rhombencephalic (posterior)
ventricles, which are connected by interventricular ducts. From the
surface geometry, we constructed a volumetric mesh of the ventricles
by initially using fTetWild~\cite{Hu2020FastWild} and subsequently
local refinement functionality in DOLFINx~\cite{dolfinx2023preprint}. 
The geometry spans approximately 600 microns along the rostrocaudal axis
and around 300 microns in the lateral direction. At most around
100 microns separates the ventral and dorsal ventricular walls.
The standard mesh consists of 141 512 tetrahedral cells with a
maximal (minimal) edge length of 14.9 \textmu
m (3.1 \textmu m) (\Crefpanels{fig:fig1}{d}). Parts of the
ventricular surfaces were marked to be lined with motile cilia
(\Crefpanels{fig:fig1}{d}, magenta
markers)~\cite{Olstad2019CiliaryDevelopment}.

@@

\subsection{Computational CSF dynamics in the brain ventricles}
The beating motion of the cilia generates CSF flow with persistent rotational
structures~\cite{Olstad2019CiliaryDevelopment}
within the brain ventricles (\Crefpanels{fig:fig1}{a--c}).
In addition, cardiac pulsations induce pulsatile CSF flow.
We model this flow of CSF by the time-dependent,
incompressible Stokes equations which read as follows.
Find the CSF velocity $\uu = \uu(\xx, t) = (u_x(\xx, t), u_y(\xx, t), u_z(\xx, t))$
and the CSF pressure $p = p(\xx, t)$ for $\xx \in \Omega$ and time $t>0$, such that
\begin{subequations}
    \begin{align}
      \rho \pdifft{\uu} - \nabla \cdot \bsig &= \mathbf{0}
      \quad \text{in } \Omega,\\
      \nabla \cdot \uu &= 0 
      \quad  \text{in } \Omega,\label{eq:stokes_eqs_cont} 
  \end{align}
  \label{eq:stokes_eqs}% 
\end{subequations}%
where the stress tensor is defined as $\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$
and $\beps(\uu) = \frac{1}{2}\left(\nabla \uu + (\nabla\uu)^T\right)$ is the strain-rate tensor. 
Bold-face characters denote vectors or tensors, and $\mathbf{I}$ is the identity tensor in three dimensions.
Furthermore, $\rho=1000 \ \mathrm{kg/m^3}$ is the CSF density
and $\mu=0.7 \ \mathrm{mPa\cdot s}$ is the dynamic viscosity
of the CSF~\cite{Bloomfield1998EffectsFluid}.
We introduce the tangential traction $\bsigpar$ as the
tangential component of the traction $\hat{\bsig}=\bsig\nn$:
\begin{equation*}
    \bsigpar = P_{\nn}(\hat{\bsig}) = (\mathbf{I} - \nn\otimes\nn)\hat{\bsig},
\end{equation*}
where $\nn$ is the outer unit normal of the boundary surface $\Gamma = \partial\Omega$,
and $P_{\nn}(\rr)$ is the tangential projection of a vector $\rr$ onto $\Gamma$.
We consider three types of boundary conditions:
(i) \emph{tangential traction} (no flow normal to the boundary
$\uu \cdot \nn = 0$ with a stress $\btau$ applied tangentially on the boundary $\bsigpar = \btau$),
(ii) \emph{normal pressure} (a pressure $\Tilde{p}$ applied in
the normal direction: $(\mu\nabla\uu - p\mathbf{I})\nn = \Tilde{p}(t)\nn$),
and (iii) \emph{free slip} (no flow normal to the boundary
$\uu \cdot \nn = 0$ and no tangential traction $\bsigpar = {\bf 0}$).
We assume that the system starts at rest: $\uu(\xx, t=0) = \mathbf{0}$.

To study the impact of the motile cilia and the cardiac pulsations on the CSF flow,
we consider the following computational flow scenarios,
each predicting a CSF flow velocity $\uu$ and pressure $p$.
In all of the scenarios, $\Gs$ denotes a free-slip boundary.
\begin{itemize}
    \item A \emph{baseline} flow model representing a wild-type zebrafish including 
    contributions both from the cilia forces and the cardiac pulsations.
    In this scenario, we let $\Gamma=\Gs\cup\Gc\cup\Gp$ and apply the tangential traction $\btau$ on the cilia boundary
    $\Gc$ and prescribe the normal pressure $\Tilde{p}$ on the anterior/posterior boundary $\Gp$. % Model 0/Model C

    \item A \emph{cilia-driven/no-cardiac} flow model including contributions from the
    cilia forces but no cardiac pulsations.
    Here, $\Gamma=\Gs\cup\Gc$, and we apply a tangential traction $\btau$ on the cilia boundary $\Gc$.
    In the absence of cardiac-induced flow, the anterior/posterior boundary $\Gp=\emptyset$. % Model I/Model A

    \item A \emph{cardiac-induced/no-cilia} flow model including contributions from the
    cardiac pulsations but no cilia forces.
    In this scenario, we disregard the cilia and consider $\Gamma=\Gs\cup\Gp$, prescribing the normal pressure $\Tilde{p}$ on the
    anterior/posterior boundary $\Gp$.  % Model II/Model B
\end{itemize}
The tangential traction and normal pressure are described further below.

@@

\subsection{Computational model of solute transport after photoconversion}
We simulate transport of photoconverted proteins within the brain ventricles by
modeling transport of a solute with concentration $c(\xx, t)$ for $\xx\in\Omega$ and
time $t > 0$ by the advection-diffusion equation
\begin{equation}
    \pdifft{c} + \nabla\cdot\JJ = 0 \quad \text{in } \Omega\setminus\Omega_{\mathrm{pc}},
    \label{eq:adv_diff_strong}
\end{equation}
where the total concentration flux $\JJ$ is assumed to consist of an advective and a diffusive flux:
\begin{equation*}
    \JJ = c\uu - D\nabla c .
\end{equation*}
Here, $\uu$ is the CSF velocity field, governed by the Stokes equations (\Cref{eq:stokes_eqs}),
and $D$ is the molecular diffusion coefficient of the solute. 
The ventricular walls are assumed to be impermeable (admitting \emph{no flux}: $\JJ\cdot\nn=0$),
except at the anterior/posterior boundary $\Gp$.
At the anterior/posterior boundary, we permit the solute concentration to be advected into
and out of the domain in a manner that ensures conservation of mass
(see~\Cref{subsec:adv_diff_eq_weak_form} for details).

@@

The numerical solution schemes for the Stokes equations and the
advection-diffusion equation were implemented and solved numerically with
DOLFINx~\cite{dolfinx2023preprint} with petsc4py~\cite{Dalcin2011ParallelPython}
as linear algebra backend. The simulation code is openly
available\fixme{~\cite{zenodo-link-to-all-code}}.
A direct solver using MUMPS~\cite{Amestoy2011Mumps} is used to solve the
linear systems resulting from the discretized Stokes equations.
For the discretized advection-diffusion equation, we use a flexible GMRES solver~\cite{fgmres}
with block Jacobi preconditioning~\cite{Jacobi1845UeberGleichungen} for the outer solves
and ILU preconditioning for the inner solves~\cite{chan1997approximate}. 
We discretize the time-dependent diffusion equation (without advection)
and use the assembled linear system matrix
as the preconditioner matrix for the advection-diffusion problem.
Numerical experiments verified the accuracy of the model implementation
(\Cref{sec:appB_verification}).
